---
title: "income_data_handling"
output: github_document
---
```{r}
library(tidyverse)
```



```{r add data source}
pums_raw = read.csv("./data/selected_pums.csv")

zcta_to_puma = readxl::read_xls("./data/zcta10_to_puma10_nyc.xls", sheet = 2) %>% 
  select(zcta = zcta10, puma = puma10) %>% 
  mutate(puma = as.numeric(puma))

zip_to_zcta = readxl::read_xls("./data/zip_to_zcta10_nyc.xls", sheet = 2) %>% 
  select(zipcode, zcta = zcta5)
```

```{r pums data wrangling}
pums_data = 
  pums_raw %>% 
  select(puma = PUMA10, income = PINCP, year = ADJINC) %>% 
  filter(puma != -9) 

pums_data$year = recode(pums_data$year, 
                        "1042852" = "2012",
                        "1025215" = "2013",  
                        "1009585" = "2014", 
                        "1001264" = "2015")

```


## Current Problems
tranfer PUMA10 code into locations.(Ref. found)  
  
match PUMA names with the location in the HIV dataset.  
  
cannot find name files for PUMA00.  
  
## Column names meaning  
  
This data is selected based on the variables that we may need
  
PUMA00 --  
Public use microdata area code 2000, which is the area code used before 2012.  
-9 means this classifications is N/A for data collected after 2012. 
  
PUMA10 --  
Public use microdata area code 2010, which is the area code used after 2012.  
-9 means this classifications is N/A for data collected prior to 2012.   
  
ADJINC --  
Adjustment factor for income and earnings dollar amounts  
1073094 -- 2011 factor  
1042852 -- 2012 factor  
1025215 -- 2013 factor  
1009585 -- 2014 factor  
1001264 -- 2015 factor  
  
PINCP --  
Total persons's income  
  

RAC3P05 --  
Recoded detailed race code for data collected prior to 2012  
  
RAC3P12 --  
Recoded detailed race code for data collected in 2012 or later  
  
For more info, you can check the data dictionarty "https://www2.census.gov/programs-surveys/acs/tech_docs/pums/data_dict/PUMS_Data_Dictionary_2011-2015.pdf?#"
  
Data from The ACS Public Use Microdata Sample files (PUMS), u can find it here "https://factfinder.census.gov/faces/tableservices/jsf/pages/productview.xhtml?pid=ACS_pums_csv_2011_2015&prodType=document"

This is the ZCTA10 to PUMA10 file
"http://faculty.baruch.cuny.edu/geoportal/resources/nyc_geog/nyc_puma_neighborhood.xls"

This thi the ZIP code to ZCTA10 file
"http://faculty.baruch.cuny.edu/geoportal/resources/nyc_geog/zip_to_zcta10_nyc_revised.xls"