---
title: "Regression"
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
library(tidyverse)
library(plotly)
library(rvest)
library(httr)
library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
theme_set(theme_bw() + theme(legend.position = "bottom") + theme(plot.title = element_text(hjust = 0.5)))
```

```{r import_data, message=FALSE, echo = FALSE}
id = "18ouzlI-Hs3t_DJW4rGUQwvHS0BurTmCV"
UHF_zipcode = 
  read_csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id)) %>% 
  slice(-43) %>% 
  select(-Borough) %>% 
  rename("UHF" = "UHF Neighborhood") %>% 
  janitor::clean_names()
```

```{r combine_data, warning=FALSE, message=FALSE, echo = FALSE}
raw_hiv = 
  GET("https://data.cityofnewyork.us/api/views/fju2-rdad/rows.csv") %>% 
  content("parsed") %>% 
  janitor::clean_names()

combine_hiv = 
  right_join(UHF_zipcode, raw_hiv, by = "uhf") %>%
  janitor::clean_names() %>% 
  separate(zip_code, into = c("zipcode1", "zipcode2", "zipcode3", 
                              "zipcode4", "zipcode5", "zipcode6", "zipcode7", "zipcode8",
                              "zipcode9"), sep = ", ") %>% 
  gather(key = zip_code, value = zipcode_value, zipcode1:zipcode9) %>% 
  filter(!is.na(zipcode_value)) %>% 
  rename("zipcode" = "zipcode_value") %>% 
  select(zipcode, everything(), -zip_code)
```

```{r pin_data, message=FALSE, echo = FALSE}
r = GET('http://data.beta.nyc//dataset/3bf5fb73-edb5-4b05-bb29-7c95f4a727fc/resource/6df127b1-6d04-4bb7-b983-07402a2c3f90/download/f4129d9aa6dd4281bc98d0f701629b76nyczipcodetabulationareas.geojson')
nyc_zipcode = readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)

zipcode_lat_lng = nyc_zipcode@data %>% 
  select(zipcode = postalCode, longitude, latitude) %>% 
  mutate(zipcode = as.character(zipcode))
  
combine_hiv1 = 
  full_join(zipcode_lat_lng, combine_hiv, by = "zipcode") %>% 
  mutate(longitude = as.numeric(longitude), latitude = as.numeric(latitude)) %>% 
  group_by(uhf) %>% 
  summarise(lng = mean(longitude),
            lat = mean(latitude)) %>% 
  filter(!(uhf == "Pelhem - Throgs Neck"))
```

```{r load_income_data, warning=FALSE, message=FALSE, echo = FALSE}
id1 = "1qGkhF7qXf_Qa8YTePLgqvEYfVLWX1tFs"
pums_raw = read_csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id1))


temp = tempfile(fileext = ".xls")

dataURL = "http://faculty.baruch.cuny.edu/geoportal/resources/nyc_geog/nyc_zcta10_to_puma10.xls"

download.file(dataURL, destfile = temp, mode = 'wb')

zcta_to_puma = readxl::read_xls(temp, sheet = 2) %>% 
  select(zcta = zcta10, puma = puma10) %>% 
  mutate(puma = as.numeric(puma))

dataURL = "http://faculty.baruch.cuny.edu/geoportal/resources/nyc_geog/zip_to_zcta10_nyc_revised.xls"

download.file(dataURL, destfile = temp, mode = 'wb')
zip_to_zcta = readxl::read_xls(temp, sheet = 2) %>% 
  select(zipcode, zcta = zcta5) 
```

```{r income_data_wrangling, echo = FALSE}
pums_data = 
  pums_raw %>% 
  select(puma = PUMA10, income = PINCP, year = ADJINC) %>% 
  filter(puma != -9)  # remove data from 2011 due to lack of area information

pums_data$year = recode(pums_data$year, 
                        "1042852" = "2012",
                        "1025215" = "2013",  
                        "1009585" = "2014", 
                        "1001264" = "2015")   

pums_data = 
  pums_data %>% 
  group_by(year, puma) %>% 
  summarise(mid_income = median(income, na.rm = TRUE)) %>% 
  ungroup()         # calculate yearly median income for each area
```

```{r zipcode match, echo = FALSE}
puma_to_zipcode = right_join(zip_to_zcta, zcta_to_puma, by = "zcta") %>%   # generaate a puma to zipcode file
  select(puma, zipcode)

income_zipcode = right_join(pums_data, puma_to_zipcode, by = "puma") %>%  # matching zipcode with median income data
  select(year, zipcode, mid_income) %>% 
  mutate(year = as.numeric(year))

combine_hiv_income = 
  left_join(combine_hiv, income_zipcode, by = c("year", "zipcode"))

income_hiv = combine_hiv_income
```

## HIV diagnoses regression

### HIV diagnoses

```{r echo = FALSE}
income_hiv %>% 
  filter(year != "2011" & age != "All") %>%
  lm(hiv_diagnoses ~ borough + gender + age + mid_income, data = .) %>% 
  summary() %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)

income_hiv %>% 
  filter(year != "2011" & race != "All") %>%
  mutate(race = fct_relevel(race, "White")) %>% 
  lm(hiv_diagnoses ~ borough + gender + race + mid_income, data = .) %>% 
  summary() %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

**Interpretation**:
HIV incidence among male are significantly higher than that among female, which can be explained by gender differences with respect to HIV/AIDS depend on patterns of disease transmission. HIV incidence in adults aged 20 to 29, 30 to 39, 40 to 49 and 50 to 59 are all significantly higher than young people aged 13 to 19. HIV incidence in Black and Latino/Hispanic population are significantly higher than that in white population. The effect of income is not significant.

### HIV related death rate

```{r echo = FALSE}
income_hiv %>% 
  filter(year != "2011" & age != "All" & hiv_related_death_rate != 99999) %>%
  lm(hiv_related_death_rate ~ borough + gender + age + mid_income, data = .) %>% 
  summary() %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)

income_hiv %>% 
  filter(year != "2011" & race != "All" & hiv_related_death_rate != 99999) %>%
  mutate(race = fct_relevel(race, "White")) %>%
  lm(hiv_related_death_rate ~ borough + gender + race + mid_income, data = .) %>% 
  summary() %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

**Interpretation**:
HIV related death rate among male are lower than that among female, which is adverse to the effect of sex on HIV diagnoses. This result indicates that the restricted social autonomy of women and young girls can reduce their ability to access sexual health and HIV services. HIV related death rate in adults aged 20 or older are significantly higher than young people aged 13 to 19, and this rate increased as the increase of age. HIV related death rate in Latino/Hispanic population are significantly higher than that in white population, which may be explained by unequal access to HIV treatments. The effect of income is not significant.