---
title: "Visualization"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
theme_set(theme_bw() + theme(legend.position = "bottom") + theme(plot.title = element_text(hjust = 0.5)))
library(plotly)
```

First, import and tidy data:

```{r message=FALSE, echo = FALSE}
hiv_data = read_csv("./data/DOHMH_HIV_AIDS_Annual_Report.csv") %>% 
  janitor::clean_names() %>% 
  rename(neighborhood = uhf) %>% 
  filter(year != "ALL", borough != "All", neighborhood != "All", gender != "All") %>% 
  mutate(year = as.character(year), age = as.factor(age))

hiv_data2 = read_csv("./data/DOHMH_HIV_AIDS_Annual_Report.csv") %>% 
  janitor::clean_names() %>% 
  rename(neighborhood = uhf) %>% 
  mutate(year = as.character(year), age = as.factor(age))
```

gender neighborhood VS hiv

```{r fig.width = 8, fig.height = 6}
neb_plot = hiv_data %>% 
  group_by(neighborhood, gender) %>% 
  filter(age != "All") %>%
  summarise(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(x = reorder(neighborhood, sum_hiv), y = sum_hiv, color = gender)) + 
  coord_flip() +
  geom_point() +
  labs(
        title = "Gender and Neighborhood Influence on HIV Incidence",
        x = "Neighborhood",
        y = "HIV diagnoses",
        caption = "Data from the ..."
      )

ggplotly(neb_plot)
```

The number of HIV diagnoses is higher among male than female in all neighborhoods. Beford Stuyvesant - Crown Heights have the most HIV diagnoses for both men and women.

gender, age vs hiv

```{r fig.width = 8, fig.height = 6}
age_plot = hiv_data %>% 
  filter(age != "All") %>% 
  group_by(gender, age) %>% 
  summarise(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(y = sum_hiv, x = age, fill = gender)) + 
  geom_bar(stat = "identity", alpha = 0.8, position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
        title = "Gender and Age Influence on HIV Incidence",
        x = "Age range",
        y = "HIV diagnoses",
        caption = "Data from the ..."
      ) 

ggplotly(age_plot)

age_plot2 = hiv_data2 %>% 
  filter(borough == "All", neighborhood == "All", race == "All",  gender != "All", age != "All") %>% 
  group_by(gender, age) %>% 
  summarise(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(y = sum_hiv, x = age, fill = gender)) + 
  geom_bar(stat = "identity", alpha = 0.8, position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2") +
  labs(
        title = "Gender and Age Influence on HIV Incidence",
        x = "Age range",
        y = "HIV diagnoses",
        caption = "Data from the ..."
      ) 

ggplotly(age_plot2)
```

gender race vs hiv

```{r fig.width = 8, fig.height = 6}
race_plot = hiv_data %>% 
  filter(race != "All") %>% 
  group_by(gender, race) %>% 
  summarise(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(y = sum_hiv, x = reorder(race, sum_hiv), fill = gender)) + 
  geom_bar(stat = "identity", alpha = 0.8, position = position_dodge()) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(
        title = "Race and Gender Influence on HIV Incidence",
        x = "Race",
        y = "HIV diagnoses",
        caption = "Data from the ..."
      ) 

ggplotly(race_plot)

race_plot2 = hiv_data2 %>% 
  filter(borough == "All", neighborhood == "All", race != "All",  gender != "All", age == "All") %>% 
  group_by(gender, race) %>% 
  summarise(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(y = sum_hiv, x = reorder(race, sum_hiv), fill = gender)) + 
  geom_bar(stat = "identity", alpha = 0.8, position = position_dodge()) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(
        title = "Race and Gender Influence on HIV Incidence",
        x = "Race",
        y = "HIV diagnoses",
        caption = "Data from the ..."
      ) 

ggplotly(race_plot2)
```

hiv diagnoses in borough with most hiv over years

```{r} 
hiv_data2 %>%
  filter(borough != "All", neighborhood == "All", gender == "All", age == "All", race == "All") %>% 
  group_by(borough) %>% 
  summarize(sum_hiv = sum(hiv_diagnoses)) %>% 
  arrange(desc(sum_hiv))

year_plot = hiv_data %>% 
  mutate(year = as.integer(year)) %>% 
  filter(borough == "Brooklyn" & gender == "Male" & age == "20 - 29") %>% 
  group_by(year, neighborhood) %>% 
  summarize(sum_hiv = sum(hiv_diagnoses)) %>% 
  ggplot(aes(x = year, y = sum_hiv, color = neighborhood)) +
  geom_line()
ggplotly(year_plot)
```

Limitations:

We can not visualize the effect of age and race at the same time.